\documentclass[10pt,a4paper,roman]{moderncv}        
\moderncvstyle{classic}                            
\moderncvcolor{green}
\usepackage[utf8]{inputenc}
\usepackage[scale=0.75]{geometry}
\usepackage[spanish,english,es-lcroman]{babel}
\usepackage{enumerate}
% personal data
\name{Ali Gharouni}{\\ \normalfont Postdoctoral Fellow}
\title{Cover Letter} 
\address{Department of Mathematics and Statistics}{McMaster University, Hamilton, Canada}
\email{aligharouni@gmail.com}  


\begin{document}
\recipient{\leavevmode}{ }

\date{\today}
\opening{Dear Editor,}

\closing{Please find enclosed our Matters Arising article. Thank you for considering this submission. Should you have any questions, please do not hesitate to correspond with me at aligharouni@gmail.com. I look forward to hearing from you.\\
\leavevmode
\\
Sincerely,}
\makelettertitle

% \noindent 
On behalf of my co-author, I am submitting our manuscript entitled "Comment on Juul et al. (2021) (?)" for exclusive consideration for publication in Nature Physics. This is a minor commentary on the paper entitled: "Fixed-time descriptive statistics underestimate extremes of epidemic curve ensembles", Juul et al. 2021 (Nature Physics 17 (1): 5-8 doi: 10.1038/s41567-020-01121-y). It is notable we have contacted the authors and sent them a copy of our commentary. The authors have responded to us and they seem happy with our comments. Also we have not had any prior discussions with a Nature Physics editor about the work described in the manuscript. 

More specifically, our research article accomplishes the following.
we provide alternative, and more principled, curved-based statistics approaches to approximate the most central set which represents the central 50\% of the ensemble. In particular, we use two functional ranking methods; (1) the sampling-based, fast and robust functional boxplot, and (2) a multivariate generalization of ranking the curves by using Mahalanobis distance among features of interest. We apply our methods on the dataset of the original paper and compare our results with theirs.

Possible reviewers for our paper include
\begin{enumerate}
\item ? maybe Carl T. Bergstrom (cbergst@uw.edu)
\item ?
\item ?
\end{enumerate}

Coauthors of the present work include
\begin{enumerate}
\item Benjamin M. Bolker \\
Department of Mathematics \& Statistics,\\
McMaster University, Hamilton, Canada\\
Email: bolker@mcmaster.ca
\end{enumerate}

\vspace{0.5cm}

\makeletterclosing

\end{document}