---
title: "conj_example"
author: "Ali"
date: "`r format(Sys.time(),'%d %b %Y')`"
output: html_document
bibliography: ../AliMac.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Under construction**

```{r source}
source('functions.R')
```

## conjecture


**conjecture**: if we have a sufficiently sensible space (metric space?), we can define the *centroid* of a set of points, which may not be a member of the set. Then for a rank $r$, the (1) set of $r$ closest points to the centroid is the same as (2) the set of $r$ points with the minimum average distance to all of the other points. ? (Right now we're using the second characteristic to define the set of central points, and we're not computing the centroid - we're computing the 'most central' point in the set [i.e. the one with minimum distance to all other points, probably == the closest to the centroid])

## example

One potential problem with Mahalanobis distances: if (for example) the probe distribution is
strongly bimodal, then scaling factors/correlations derived from the overall data set may
not be appropriate for scaling the components of distance between two trajectories whose features
put them in the same mode/component of the distribution ...

```{r bimodal}
set.seed(101)
m1 <- MASS::mvrnorm(50, mu=c(6,6), Sigma=matrix(c(1,-0.7,-0.7,1),2))
m2 <- MASS::mvrnorm(50, mu=c(-6,-6), Sigma=matrix(c(1,-0.7,-0.7,1),2))
m <- rbind(m1,m2)
plot(m[,1],m[,2])
```


## Euclidean Norm; the case that the conjecture is held
```{r counter_eg}
## set of points
x <- c(1:3)
n <- length(x)

## Euclidean norm, where 
mat1  <- as.matrix(dist(as.matrix(x), method = "euclidean", diag = TRUE, upper = TRUE))
eucl1 <- colMeans(mat1)

eucl2 <- apply(as.matrix(x),1,FUN=function(y){dist(c(y,mean(x)))} )

rank(eucl1)
rank(eucl2)

```

## Log Euclidean Norm; the case that the conjecture is not hold

Following @arsigny2006log, the definition of log Euclidean norm is understood for Real numbers as follows. 

```{r logeuclid}
 x <- c(1:10)
 n<- length(x)
mat<- matrix(NA,n,n)

diag(mat)<-0 

for (i in 1:(n-1) ){
  for(j in (i+1):n){
      mat[i,j] <- logedu(x[i],x[j])
  }
}

mat <- make_symm(mat)
y1 <- colMeans(mat)
rank(y1)
  
m <- mean(x)
y2 <- apply(matrix(x),1,FUN=function(y){logedu(y,m)})
rank(y2) 
```



# References





